Simulating High-Dimensional Multivariate Data using the \pkg{bigsimr} Package

### 2021-06-01 030-bigsimr-package.Rmd
idk
Revise for JSS submission pre cross ref fix

# The `bigsimr` R package {#package}

@@

This section describes a low-dimensional (2D) random vector simulation workflow via the `bigsimr` `R` package (https://github.com/SchisslerGroup/r-bigsimr). The package `bigsimr` provides an interface to the native code written in Julia, registered as the `Bigsimr` Julia package (https://github.com/adknudson/Bigsimr.jl). In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `bigsimr` (https://github.com/SchisslerGroup/python-bigsimr/) that interfaces with the Julia `Bigsimr` package. The Julia package provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).

@@

## Bivariate example description

We illustrate the use of `bigsimr` motivated by the New York air quality data set (`airquality`) included in the R `datasets` package. First, we load the `bigsimr` library and a few other convenient data science packages, including the syntactically-elegant `tidyverse` suite of `R` packages. The code chunk below prepares the computing environment:

@@

For simplicity and to provide a minimal working example, we consider bivariate simulation of the two `airquality` variables `Temperature`, in degrees Fahrenheit, and `Ozone` level, in parts per billion.

@@

Figure \@ref(fig:ch030-aq-joint-dist) visualizes the bivariate relationship between Ozone and Temperature. We aim to simulate random two-component vectors mimicking this structure. The margins are not normally distributed; particularly the Ozone level exhibits a strong positive skew.

```{r ch030-aq-joint-dist, cache=FALSE, fig.width= 8, fig.cap="Bivariate scatterplot of Ozone vs. Temp with estimated marginal densities. We model the Ozone data as marginally log-normal and the Temperature data as normal."}
p0 <- ggplot(df, aes(Temp, Ozone)) +
  geom_point(size = 1) +
  theme(legend.position = "none") + 
  labs(x = "Temperature")

@@

Next, we specify the marginal distributions and correlation coefficient (both type and magnitude). Here the analyst is free to be creative. For this example, we avoid goodness-of-fit considerations to determine the marginal distributions. But it seems sensible without domain knowledge to estimate these quantities from the data, and `bigsimr` contains fast functions designed for this task.

@@

Based on the estimated densities in Figure \@ref(fig:ch030-aq-joint-dist), we assume `Temp` is normally distributed and `Ozone` is log-normally distributed, as the latter values are positive and skewed. We use the well-known, unbiased estimators for the normal distribution's parameters and maximum likelihood estimators for the log-normal parameters:

@@

Next, we configure the input marginals for later input into `rvec`. The marginal distributions are specified using `Julia`'s `Distributions` package and stored in a vector.

@@

As mentioned, the user must decide how to describe correlation based on the particulars of the problem. For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$. We also support Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance). These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

@@

First we use `cor_bounds` to ensure that the pairwise target correlation values are valid prior to mapping step which constructs $R_X$ for the MVN input into NORTA (see Section \@ref(algorithms)). `cor_bounds` estimates the pairwise lower and upper correlation bounds using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(margins)
bounds$lower
bounds$upper
```

Since our single estimated Spearman correlation is within the theoretical bounds, the correlation is valid as input to `rvec`. But even if the 2-dimensional bounds are satisfied for each pair of margins, this does not guarantee the feasibility of a $d-$variate distribution [@BF17].

To provide higher dimensional feasibility/admissibility checking, we begin by mapping using `cor_convert` and check admissibility.

```{r ch030-cor-convert, echo=TRUE}
# Step 1. Mapping
(R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson))
# Step 2. Check admissibility
bs$iscorrelation(R_X)
```

The bounds on the Pearson correlation coefficient $R_P$ between these margins are restricted as seen below in our MC estimated correlation bounds:

@@

Our MC estimate of the bounds slightly underestimates the theoretic bounds of $(-0.881, 0.881)$. An analytic derivation is presented as an Appendix.

@@

Finally, we arrive at the main function of `bigsimr`: `rvec`. We now simulate $B=10,000$ random vectors from the assumed joint distribution of Ozone levels and Temp.

>>>>

# The \pkg{bigsimr} R package {short-title="bigsimr-pkg" #package}

@@

This section describes a LD random vector simulation workflow via the `bigsimr` `R` package. Our `R` package provides an interface to the native code written in Julia (registered as the `Bigsimr` Julia package. In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `bigsimr` that interfaces with the Julia `Bigsimr` package. The Julia package provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).

@@

## Bivariate example

We illustrate the use of `bigsimr` using the New York air quality data set (`airquality`) included in the R `datasets` package. First, we load the `bigsimr` library and a few other convenient data science packages, including the syntactically-elegant `tidyverse` suite of `R` packages. The code chunk below prepares the computing environment:

@@

Below, we describe a minimal working example --- a bivariate simulation of two airquality variables: `Temperature`, in degrees Fahrenheit, and `Ozone` level, in parts per billion.

@@

Figure \@ref(fig:ch030-aq-joint-dist) visualizes the bivariate relationship between `Ozone` and `Temperature`. We aim to simulate random two-component vectors mimicking this structure. The margins are not normally distributed; `Ozone` level exhibits a strong positive skew.

```{r ch030-aq-joint-dist, cache=FALSE, fig.width= 8, fig.cap="Bivariate scatterplot of Ozone versus Temp with estimated marginal densities. The Ozone data are modeled marginally as log-normal and the Temperature data as normal."}
p0 <- ggplot(df, aes(Temp, Ozone)) +
  geom_point(size = 1) +
  theme(legend.position = "none") + 
  labs(x = "Temperature")

@@

Next, we specify the marginal distributions and correlation coefficient (both type and magnitude). Here the analyst is free to be creative. For this example, we avoid goodness-of-fit considerations to determine the marginal distributions. It is sensible without domain knowledge to estimate these quantities from the data, and `bigsimr` contains fast functions designed for this task.

@@

Based on the estimated densities in Figure \@ref(fig:ch030-aq-joint-dist), we assume `Temp` is normally distributed and `Ozone` is log-normally distributed, as the latter values are positive and skewed. We use the classical unbiased estimators for the normal distribution's parameters and maximum likelihood estimators for the log-normal parameters:

@@

Now, we configure the input marginals for later input into `rvec`. The marginal distributions are specified using `Julia`'s `Distributions` package and stored in a vector.

@@

As mentioned, the user must decide how to describe correlation based on the particulars of the problem. For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$. We also support Pearson correlation coefficient matching, while cautioning the user to check the performance for the distribution at hand (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance). Note that hese estimation methods are classical approaches, not specifically designed for high-dimensional correlation estimation.

@@

Once a target correlation matrix is specified, first convert the dependency type to Pearson to correctly specify the MVN inputs. Then, check that the converted matrix $R_X$ is a valid correlation matrix (PD and 1s on the diagonal). For this bivariate example, $R_X$ is a valid correlation matrix. Typically in HD the resultant MVN Pearson correlation matrix is indefinite and requires approximation (see the HD examples in subsequent sections).

```{r ch030-cor-convert, echo=TRUE}
# Step 1. Mapping
(R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson))
```

```{r ch030-cor-check, echo=TRUE}
# Step 2. Check admissibility
bs$iscorrelation(R_X)
```

Despite being a valid correlation matrix under this definition, marginal distributions induce Frechet bounds on the possible Pearson correlation values. To check that the converted correlation values fall within these bounds, use `bigsimr::cor_bounds` to estimate the pairwise lower and upper correlation bounds. `cor_bounds` uses the Generate, Sort, and Correlate algorithm of @DH2011. For the assumed marginals, the Pearson correlation coefficient is not free to vary in [-1,1], as seen below. Our MC estimate of the bounds slightly underestimates the theoretic bounds of $(-0.881, 0.881)$. (See an analytic derivation presented in the Appendix). Since our single Pearson correlation coefficient is within the theoretical bounds, the correlation matrix is valid for our simulation strategy.

@@

@@

A single `rvec` execution simulates the desired $10,000$ random vectors from the assumed bivariate distribution of `Ozone` and `Temp`. Note that the target input in specified using the pre-computed MVN correlation matrix.

### 2021-06-01 040-monte-carlo-evaluation.Rmd
Revise for JSS submission pre cross ref fix

Before applying our methodology to real data simulation, we conduct several Monte Carlo studies to investigate method performance. Since marginal parameter matching in our scheme is essentially a sequence of univariate inverse probability transforms, the challenging aspects are the accuracy of dependency matching and computational efficiency at high dimensions. To evaluate our methods in those respects, we design the following numerical experiments to first assess accuracy of matching dependency parameters in bivariate simulations and then time the procedure in increasingly large dimension $d$.

@@

We select bivariate simulation configurations to ultimately simulate our motivating discrete-valued RNA-seq example, and, so we proceed in increasing complexity, leading to the model in our motivating application in Section \@ref(examples). We begin with empirically evaluating the dependency matching across all three supported correlations --- Pearson's, Spearman's, and Kendall's --- in identical, bivariate marginal configurations. For each pair of identical margins, we vary the target correlation across $\Omega$, the set of possible admissible values for correlation type, to evaluate the simulation's ability to obtain the theoretic bounds. The simulations progress from bivariate normal, to bivariate gamma (non-normal yet continuous), and bivariate negative binomial (mimicking RNA-seq counts).

Table \@ref(tab:sims) lists our identical-marginal, bivariate simulation configurations. We increase the simulate replicates $B$ to check that our results converge to the target correlations and gauge statistical efficiency. We select distributions beginning with a standard multivariate normal (MVN) as we expect the performance to be exact (up to MC error) for all correlation types. Then, we select a non-symmetric continuous distribution: a standard (rate =1) two-component multivariate gamma (MVG). Finally, we select distributions and marginal parameter values that are motivated by our RNA-seq data, namely values proximal to probabilities and sizes estimated from the data (see [Example applications](examples) for estimation details). Thus we arrive at a multivariate negative binomial (MVNB) $p_1 = p_2 = 3\times10^{-4}, r_1 = r_2 = 4, \rho \in \Omega$.

Table: (\#tab:sims) Identical margin, bivariate simulation configurations to evaluate correlation matching accuracy and efficiency.

\begin{table}[]
\begin{tabular}{@{}lcr@{}}
\toprule
Simulation Reps $B$ & Correlation Types & Identical-margin 2D distribution \\ \midrule
$1,000$ & Pearson ($\rho_P$) & ${\bf Y} \sim MVN( \mu= 0 , \sigma = 1, \rho_i ), i=1,\ldots,100$ \\
$10,000$ & Spearman ($\rho_S$) & ${\bf Y} \sim MVG( shape = 10, rate = 1, \rho_i ), i=1,\ldots,100$ \\
$100,000$ & Kendall ($\tau$) & ${\bf Y} \sim MVNB(p = 3\times10^{-4}, r = 4,\rho_i), i=1,\ldots,100$ \\ \bottomrule
\end{tabular}
\end{table}

For each of the unique 9 simulation configurations described above, we estimate the correlation bounds and vary the correlations along a sequence of 100 points evenly placed within the bounds, aiming to explore $\Omega$. Specifically, we set correlations $\{ \rho_1 = ( \hat{l} + \epsilon), \rho_2 = (\hat{l} + \epsilon) + \delta, \ldots, \rho_{100} = (\hat{u} - \epsilon) \}$, with $\hat{l}$ and $\hat{u}$ being the estimated lower and upper bounds, respectively, and increment value $\delta$. The adjustment factor, $\epsilon=0.01$, is introduced to handle numeric issues when the bound is specified exactly.

@@

Overall, the studies show that our methodology is generally accurate across the entire range of possible correlation values all three dependency measures, at least in these limited simulation settings for the rank-based correlations. Our Pearson matching performs nearly as well as Spearman or Kendall, except for a slight increase in error for negative binomial case. This is due the particularly large counts generated with our choice of parameters for $p$ and $r$.

@@

With information of our method's accuracy from a low-dimensional perspective, we now turn to assessing whether `bigsimr` can scale to larger dimensional problems with practical computation times. Specifically, we ultimately generate $B=1,000$ random vectors for $d=\{100, 250, 500, 1000, 2500, 5000, 10000\}$ for each correlation type, $\{Pearson, Spearman, Kendall\}$ while timing the algorithm's major steps.

To mimic the workflow, we first produce a synthetic "data set" by completing the following steps:

\setstretch{1.5}
1. Produce heterogeneous gamma marginals by randomly selecting the $j^{th}$ gamma shape parameter from $U_j \sim uniform(1,10), j=1,\ldots,d$ and the $j^{th}$ rate parameter from $V_j \sim exp(1/5), j=1,\ldots,d$, with the constant parameters determined arbitrarily.
2. Produce a random full-rank Pearson correlation matrix via `cor_randPD` of size $d \times d$.
3. Simulate a "data set" of $1,000 \times d$ random vectors via `rvec` (without matching the Pearson exactly).
\setstretch{2.0}

With the synthetic data set in hand, we complete and time the following 4 steps:

\setstretch{1.5}
1. a) Estimate the correlation matrix from the "data" in the *Compute Correlation* step.
1. b) Map the correlations to initialize the algorithm (Pearson to Pearson, Spearman to Pearson, or Kendall to Pearson) in the *Adjust Correlation* step.
2. Check whether the mapping produces a valid correlation matrix and, if not, find the nearest PD correlation matrix in the *Check Admissibility* step.
3. Simulate $1,000$ vectors via the Inverse Transform method in the *Simulate Data* step.
\setstretch{2.0}

The experiments are conducted on a MacBook Pro carrying a 2.4 GHz 8-Core Intel Core i9 processor, with all 16 threads employed during computation. The results for $d \leq 500$ are fast with all times (except Kendall) under 3 seconds. Table \@ref(tab:ch040-moderateDtab) displays the total computation time of the algorithms steps 1 through 3, including estimation step i.

@@

tabTimes %>%
    kable(booktabs = TRUE,
          linesep = c("", "", "\\addlinespace"), format="latex",
          col.names = c("Dimension",
                        "Correlation type",
                        "Total Time (Seconds)"),
          caption = 'Total time to produce 10,000 random vectors with a random correlation matrix and hetereogeneous gamma margins.')
```

The results with larger dimensional vectors show scalability to ultra-high dimensions for all three correlation types, although the total times do become much larger. 
Figure \@ref(fig:ch040-largeDfig) displays computation times for $d=\{1000, 2500, 5000, 10000\}$. For $d$ equal to 1000 and 2500, the total time is under a couple of minutes. At $d$ of 5000 and 10,000, Pearson correlation matching in the *Adjust Correlation* step becomes costly. Interestingly, Pearson is actually faster than Kendall for $d=10,000$ due to bottlenecks in *Compute Correlation* and *Check Admissibility*. Uniformly, matching Spearman correlations is faster, with total times under 5 minutes for $d=10,000$, making Spearman correlations the most computationally-friendly dependency type. With this in mind, we scaled the simulation to $d=20,000$ for the Spearman type and obtained the 1,000 vectors in under an hour (data not shown). In principle, this would enable the simulation of an entire human-derived RNA-seq data set. We note that for a given target correlation matrix and margins, steps i, 1, and 2 only need to be computed once and the third step, *Simulate Data*, is fast under all schemes for all $d$ under consideration.

```{r ch040-largeDfig, fig.cap="Computation times as d increases."}
# Set colors
numColors <- 5
numGroups <- 4
myColors <- rev(brewer.pal(n=numColors, name="Greys")[((numColors-numGroups)+1):numColors])

@@

In the bivariate studies, we chose arbitrary simulation parameters for three distributions, moving from the Gaussian to the discrete and non-normal, multivariate negative binomial. Under these conditions, the simulated random vectors sample the desired bivariate distribution across the entire range of pairwise correlations for the three dependency measures. The simulation results could differ for other choices of simulation settings. Specifying extreme correlations near the boundary or Frechet bounds could result in poor simulation performance. Fortunately, it is straightforward to evaluate simulation performance by using strategies similar to those completed above. We expect our random vector generation to perform well for the vast majority of NORTA-feasible correlation matrices, but advise to check the performance before making inferences/further analyses.

Somewhat surprising, Kendall estimation and nearest PD computation scale poorly compared to Spearman and, even, approximate Pearson matching. In our experience, Kendall computation times are sensitive to the number of cores, benefiting from multi-core parallelization. This could mitigate some of the current algorithmic/implementation shortcomings. Despite this, Kendall matching is still feasible for most high-dimensional data sets. Finally, we note that one could use our single-pass algorithm `cor_fastPD` to produce a 'close' (not nearest PD) to scale to even higher dimensions with greater loss of accuracy.

>>>>

In this section, we conduct MC studies to investigate method performance. Marginal parameter matching is straightforward in our scheme as it is a sequence of univariate inverse probability transforms. This process is both fast and exact. On the other hand, accurate and computationally efficent dependency matching presents a challenge, especially in the HD setting. To evaluate our methods in those respects, we design the following numerical experiments to first assess accuracy of matching dependency parameters in bivariate simulations and then time the procedure in increasingly large dimension $d$.

@@

We select bivariate simulation configurations to ultimately simulate our motivating discrete-valued RNA-seq example. Thus, we proceed in increasing departure from normality, leading to the multivariate negative binomial (MVNB) model for our motivating data. We begin with empirically evaluating the dependency matching across all three supported correlations --- Pearson, Spearman, and Kendall --- in identical, bivariate marginal configurations. For each pair of identical margins, we vary the target correlation across $\Omega$, the set of possible admissible values for each correlation type, to evaluate the simulation's ability to simulate all possible correlations. The simulations progress from bivariate normal, to bivariate gamma (non-normal yet continuous), and bivariate negative binomial (non-normal and discrete).

Table \@ref(tab:sims) lists our identical-marginal, bivariate simulation configurations. We increase the simulate replicates $B$ to visually ensure that our results converge to the target correlations and gauge efficiency. We select distributions beginning with a standard multivariate normal (MVN) as we expect the performance to be exact (up to MC error) for all correlation types. Then, we select a non-symmetric continuous distribution: a standard (rate =1), two-component multivariate gamma (MVG). Finally, we select distributions and marginal parameter values that are motivated by our RNA-seq data, namely values proximal to probabilities and sizes estimated from the data (see [RNA-seq data application](examples) for estimation details). Specifically, we assume a MVNB with $p_1 = p_2 = 3\times10^{-4}, r_1 = r_2 = 4, \rho \in \Omega$.

<!-- 
Table: (\#tab:sims) Identical margin, bivariate simulation configurations to evaluate matching accuracy.
\begin{table}[]
\toprule
Simulation Reps $B$ & Correlation Types & Identical-margin 2D distribution \\ \midrule
$1,000$ & Pearson ($\rho_P$) & ${\bf Y} \sim MVN( \mu= 0 , \sigma = 1, \rho_i ), i=1,\ldots,100$ \\
$10,000$ & Spearman ($\rho_S$) & ${\bf Y} \sim MVG( shape = 10, rate = 1, \rho_i ), i=1,\ldots,100$ \\
$100,000$ & Kendall ($\tau$) & ${\bf Y} \sim MVNB(p = 3\times10^{-4}, r = 4,\rho_i), i=1,\ldots,100$ \\ \bottomrule
\end{tabular}
\end{table}
-->

For each of the unique 9 simulation configurations described above, we estimate the correlation bounds and vary the correlations along a sequence of 100 points evenly placed within the bounds, aiming to explore $\Omega$. Specifically, we set correlations $\{ \rho_1 = ( \hat{l} + \epsilon), \rho_2 = (\hat{l} + \epsilon) + \delta, \ldots, \rho_{100} = (\hat{u} - \epsilon) \}$, with $\hat{l}$ and $\hat{u}$ being the estimated lower and upper bounds, respectively, with increment value $\delta$. The adjustment factor, $\epsilon=0.01$, is introduced to handle numeric issues when the bound is specified exactly.

@@

Overall, the studies show that our methodology is generally accurate across the entire range of possible correlation for all three dependency measures. Our Pearson matching performs nearly as well as Spearman or Kendall, except for a slight increase in error for negative binomial case.

@@

With information of our method's accuracy from a low-dimensional perspective, we now assess whether `bigsimr` can scale to larger dimensional problems with practical computation times. We ultimately generate $B=1,000$ random vectors for $d=\{100, 250, 500, 1000, 2500, 5000, 10000\}$ for each correlation type, {Pearson, Spearman, Kendall} while timing the algorithm's major steps.

We first produce a synthetic "data set" by completing the following steps:

\setstretch{1.5}
1. Produce heterogeneous gamma marginals by randomly selecting the $j^{th}$ gamma shape parameter from $U_j \sim uniform(1,10), j=1,\ldots,d$ and the $j^{th}$ rate parameter from $V_j \sim exp(1/5), j=1,\ldots,d$, with the constant parameters determined arbitrarily.
2. Produce a random full-rank Pearson correlation matrix via `cor_randPD` of size $d \times d$.
3. Simulate a "data set" of $1,000 \times d$ random vectors via `rvec`.
\setstretch{2.0}

With the synthetic data set in hand, we complete and time the following four steps involved in a typical workflow (including correlation estimation).

\setstretch{1.5}
1. Estimate the correlation matrix from the "data" in the *Compute Correlation* step.
2. Map the correlations to initialize the algorithm (Pearson to Pearson, Spearman to Pearson, or Kendall to Pearson) in the *Adjust Correlation* step.
3. Check whether the mapping produces a valid correlation matrix and, if not, find the nearest PD correlation matrix in the *Check Admissibility* step.
4. Simulate $1,000$ vectors in the *Simulate Data* step.
\setstretch{2.0}

The experiments are conducted on a MacBook Pro carrying a 2.4 GHz 8-Core Intel Core i9 processor, with all 16 threads employed during computation. Table \@ref(tab:ch040-moderateDtab) displays the total computation time for moderate dimesions ($d \leq 500$). In every simullation setting, the 1,000 vectors are generated rapidly, executing in under 2 seconds.

@@

tabTimes %>%
    kable(booktabs = TRUE,
          linesep = c("", "", "\\addlinespace"), format="latex",
          col.names = c("Dimension",
                        "Correlation type",
                        "Total Time (Seconds)"),
          caption = 'Total time to produce 1,000 random vectors with a random correlation matrix and hetereogeneous gamma margins.')
```

The results for $d > 500$ show scalability to ultra-high dimensions for all three correlation types, although the total times do become much larger. 
Figure \@ref(fig:ch040-largeDfig) displays computation times for $d=\{1000, 2500, 5000, 10000\}$. For $d$ equal to 1000 and 2500, the total time is under a couple of minutes. At $d$ of 5000 and 10,000, Pearson correlation matching in the *Adjust Correlation* step becomes costly. Interestingly, Pearson is actually faster than Kendall for $d=10,000$ due to bottlenecks in *Compute Correlation* and *Check Admissibility*. Uniformly, matching Spearman correlations is faster, with total times under 5 minutes for $d=10,000$, making Spearman the most computationally-friendly dependency type for our package. With this in mind, we scaled the simulation to $d=20,000$ for the Spearman type and obtained the 1,000 vectors in under an hour (data not shown). In principle, this would enable the simulation of an entire human-derived RNA-seq data set. We note that for a given target correlation matrix and margins, steps 1, 2, and 3 only need to be computed once and the fourth step, *Simulate Data*, is nearly instaneous for all settings considered.

```{r ch040-largeDfig, fig.cap="Computation times for HD multivariate gamma simulation."}
# Set colors
numColors <- 5
numGroups <- 4
myColors <- rev(brewer.pal(n=numColors, name="Greys")[((numColors-numGroups)+1):numColors])

@@

In the bivariate studies, we chose arbitrary simulation parameters for three distributions, moving from the Gaussian to discrete and non-normal MVNB. Under these conditions, the simulated random vectors sample the desired bivariate distribution across the entire range of pairwise correlations for the three dependency measures. The simulation results could differ for other choices of simulation settings. Specifying extreme correlations near the boundary or Frechet bounds could result in poor simulation performance. Fortunately, it is straightforward to evaluate simulation performance by using strategies similar to those completed above. We expect our random vector generation to perform well for the vast majority of NORTA-feasible correlation matrices, but advise to check the performance before making inferences/further analyses.

Somewhat surprising, Kendall estimation and nearest PD computation scale poorly compared to Spearman and, even, approximate Pearson matching. In our experience, Kendall computation times are sensitive to the number of cores, benefiting from multi-core parallelization. This could mitigate some of the current algorithmic/implementation shortcomings. Despite this, Kendall matching is still feasible for most HD data sets. Finally, we note that one could use our single-pass algorithm `cor_fastPD` to produce a 'close' (not nearest PD) to scale to even higher dimensions with some loss of accuracy.
